# Reproducible Research: Peer Assessment 1
========================================================
## Loading and preprocessing the data

Read in file - activity monitoring data file from assignment page
```{r,echo=TRUE}
activity<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
This is a calculation of the mean total steps taken during each day

```{r,echo=TRUE}
meanstepsday<-tapply(activity$steps,activity$date, mean)
meanstepsday
```

This is a histogram showing the mean steps taken each day
```{r,echo=TRUE}
 hist(meanstepsday, main="Mean Total Steps per Day", xlab="Mean Steps")
```

This is a calculation of the total steps per day with a histogram showing the frequency of Total step counts per day
```{r, echo=TRUE}
totalstepsday<-tapply(activity$steps,activity$date, sum)
hist(totalstepsday, main="Total Steps per Day", xlab="Total Steps")
```


This is a calculation of the median steps taken each day
```{r, echo=TRUE}
medianstepsday<-tapply(activity$steps,activity$date, median)
medianstepsday
```

## What is the average daily activity pattern?

This is a calculation of of the average number of steps per 5 minute interval
```{r, echo=TRUE}
meanstepsinterval<-tapply(activity$interval, activity$steps,mean)
meanstepsinterval
plot(meanstepsinterval, type="l",main="Average Steps per Interval",ylab="Mean Steps per Interval", xlab="5 Minute Intervals")
```

Calculate maximum steps in an interval
```{r, echo=TRUE}
maxstepsinterval<-tapply(activity$interval, activity$steps,max)
max(maxstepsinterval)
```


## Imputing missing values
Calculate number of missing values - shown in this summary as the number of NA's
```{r,echo=TRUE}
summary(activity)
```

To replace NA's in the data I did the following:
I calculated the mean of "non NA steps" in the data
```{r,echo=TRUE}
meansteps<-mean(activity$steps, na.rm=TRUE)
meansteps
```

Create new data file and replace NA values with mean of steps to fill in missing data
Show sample of 500 with NA changed to 37.38
```{r,echo=TRUE}
activity2 <- activity
activity2[is.na(activity2 <- activity)] <- meansteps
head(activity2[200:700,],500)
```

Calculate mean and median for filled in data file 
```{r,echo=TRUE}
newmeanstepsday<-tapply(activity2$steps,activity2$date, mean)
newmeanstepsday
```

This is a histogram showing the mean steps taken each day for the filled in data file
```{r,echo=TRUE}
 hist(newmeanstepsday, main="Mean Total Steps per Day Filled In Data File", xlab="Mean Steps")
```

This is a calculation of the median steps taken each day in the filled in data file
```{r, echo=TRUE}
newmedianstepsday<-tapply(activity2$steps,activity2$date, median)
newmedianstepsday
```

These values do differ from the earlier part of the assignment with the missing data.

The median shows values where there were none when values were missing.

The mean values for the filled in data file are greater in frequency than with the missing data file but the contour of the histogram is similar.


## Are there differences in activity patterns between weekdays and weekends?
Find the weekday name for each date and add to datafile
```{r,echo=TRUE}
weekdays<-weekdays(as.Date(activity2$date,'%m/%d/%Y'))
activity2days<-cbind(activity2,weekdays)
```

calculate mean steps per interval per day
```{r, echo=TRUE}
dailymeansteps<-tapply(activity2days$steps, activity2days$weekdays, mean)
dailymeansteps
```
Load timeDate package
Calculate Weekend indicator
Add as column "weekend" to activity2days
```{r,echo=TRUE}
library("timeDate")
weekend<-isWeekend(as.Date(activity2$date,'%m/%d/%Y'))
activity2days<-cbind(activity2days,weekend)
activity2days[1425:1450,]
```

Calculate mean steps per day
Plot for weekend vs weekdays
```{r,echo=TRUE}
library(ggplot2)
 g<-qplot(activity2days$date, activity2days$steps, data=activity2days, color=weekend, facets=weekend~.,xlab=("Day"),ylab=("Steps"), main=("Steps per Day  Weekend vs Weekdays"))+geom_line()
g
```


 


 
   



